---
title: "CBowersoxAssign6"
author: "Cheryl Bowersox"
date: "Saturday, March 07, 2015"
output: html_document
---
# Week 5 Assignment

Cheryl Bowersox 
IS 605 - 1
3/8/15


## Problem Set 1

### Number of possible outcomes of roll die 3 times
n = 6
possible outcomes (order matters) = $6^3$ = `r 6^3` 

possible outcomes (order does not matter ) = 

$$
  n = 6\\
  r = 3\\
  t = \frac{(6+3-1)!}{3!(6-1)!}\\
  t = 56
$$


### Probablity of sum = 3 when roll dice twice

$$
  n = 6\\
  p(x) = \frac{1}{6}\\
  p(x,y) = \frac{1}{6^2} = \frac{1}{36}\\
  2*p(x,y) = \frac{1}{18}
$$

### Probablity of same birthday
The probablity that any two people out a of population r share a birthday is

p = 1 - probablity that no two people share a birthday

q = probablity no people share


For 25 people:
$$
  n = 365\\
  k = 25\\
  q = {_{25}P_{365}} \times \frac{1}{365^{25}}
$$

```{r}
cperm <- function(n,k){
  choose(n,k) * factorial(k)
  }

d <- cperm(365,25)/(365^25)
```

$$
  p =1 - `r d` = `r 1 - d`
$$


For 50 people: 
$$
  n = 365\\
  k = 50\\
    q = {_{50}P_{365}} \times \frac{1}{365^{50}}
$$
```{r}
d <- cperm(365,50)/365^50
```

$$
  p =1 - `r d` = `r 1 - d`
$$


## Problem Set 2

### read data and calculate word and word pair probablities 

```{r}
library(stringr)
library(RCurl)
library(qdap)
library(scales)

# user input

nonum <- function(x){
  x <- strsplit(x, " ")
  x <- lapply(x,replace_number)
  x <-gsub("[[:punct:]]", "", toString(unlist(x)))
  return(x)
}

singleword <- function(w,doc){
  #function takes single word and full document data and finds 
  # number of time a word is contained in the doc
  w <- nonum(tolower(gsub("[[:punct:]]", "", w)))
  cpart <- sum(str_count(doc, w)) # count number of times it appears in any form 
  return(cpart)
}

wordpairs <- function(w1, w2,doc){
  #function takes a word pair and full document data and finds 
  # number of time a word pair is contained in the doc
  w1 <- nonum(tolower(gsub("[[:punct:]]", "", w1)))
  w2 <- nonum(tolower(gsub("[[:punct:]]", "", w2)))
  s1 <- paste("\\<",w1, " ",w2 ,"\\>", sep ="")
  s3 <- paste("\\<",w1, " ",w2 ,"\\>", sep ="")
  s2 <- paste("\\<",w2, " ",w1 ,"\\>", sep ="")
  c1 <- sum(str_count(doc, s1))  #count number of full words
  c2 <- sum(str_count(doc, s2))
  return(c1 + c2)
}

#get doc
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
fileurl= "https://raw.githubusercontent.com/cherylb/CompMath/master/assign6.sample.txt"
data <- getURL(fileurl)

docTxt <- iconv(data, "UTF-8", "ASCII", sub="?")
doc  <- gsub("[\n]", "", data)
#clean up doc
doc <- tolower(gsub("[[:punct:]]", "", doc))
doc <- gsub("[^[:alnum:] ]", "", doc)


# counting before convert numbers to words to treat numbers with more than one digit
# as a single word
n <-  sum(str_count(doc, "\\w+"))  
np <-  n - 1    # of possible adjacent pairs
doc <- replace_number(doc)

findprob1 <- function(w,doc){
  # w is single word to find
  # doc is doc to examine
  p <- singleword(w,doc)
 
  if(n != 0){
    ps1 <- p/n
  }
  
  if(ps1 == 0){
    ps1 <- 0
    }else{
    ps1 <- percent(ps1/ 100)
    }

  
  #output
  print(paste("found", p,"occurances of",  w, "as whole or part of a word"))
  print(paste("probablity found as whole or part of a word: ",ps1))
}

findprob2 <- function(u,doc){
  # u is pair of words, seperated by comma
  # doc is document to examine
  
  w1 <- unlist(strsplit(u,","))[1]
  w1 <- gsub("^\\s+|\\s+$", "", w1)
  w2 <- unlist(strsplit(u,","))[2]
  w2 <- gsub("^\\s+|\\s+$", "", w2)
  
  p <- wordpairs(w1, w2,doc)
  pw = 0
  if(np != 0){
    pw <- p/np
  }
  if(pw == 0){
    pw <- 0
    }else{
    pw <- percent(pw/ 100)
    }
  
  print(paste("found", p, " occurances of the word pair (", w1, w2,")"))
  print(paste("probablity is:", pw))
}
```
Finding single word 
inmate
monkey
he
she

```{r}
findprob1("inmate", doc)
findprob1("monkey", doc)
findprob1("he",doc)
findprob1("she",doc)

```
finding pairs of words:
said he
5 monkey
need we
is not

```{r}
findprob2("said, he", doc)
findprob2("5,monkey", doc)
findprob2("need, we", doc)
findprob2("is, not", doc)
```

