# Week 2 Assignment
Cheryl Bowersox 
IS 605 - 1
2/8/15

========================================================
## Problem Set 1


In general: $ A \times A^T \ne A^T \times A$ $\forall A$
  
  (1)   If A is a n X m matrix and B is a m X p martrix then $A \times B$ will result in an n X p matrix $\forall A, B\\$
  
  (2)   If $A$ is a n X m matrix then $A^T$ is a m X n matrix
  
  (3)   $A \times A^T $ is a n X n matrix by (1)
  
  (4)   $A^T \times A$ is a m X m matrix by (1)
  
  (5)   If $n \ne m$ then $A^T \times A$ cannot equal $A times A^T$
  
  (6)   $A \times A^t \ne A^t \times A$ $\forall A$
  
  
      
For some A where A is n x n $A \times A^T = A^T \times A\\$ 

  (1)   $B = A^T\\$
  
  (2)  then $A_{ij} = B_{ji}$
  
  (3)  If A is a diagonal square matrix $A_{ij} = B_{ji} \forall i = j$ 
  $A_{ij} = 0 = B_{ji} \forall  i \ne j$
  
  (4)   $A = B = A^T$
  (5)   A \times \A^T = A \times \A = A^T \times A
        

========================================================
## Problem Set 2
```{r}
factormat <- function(A){
  U <- A
  i <- 1
  j <- 1
  k <- 1
  n <- nrow(A)
  c <- ncol(A)
  E <- diag(n)
  l = list()
   
  while (j < n){
  # is entire col zero
    if(all(U[,j] == 0)){
      j = j + 1
    }
     i = j
     
     while (i <= n ){
     
       if (U[i,j] == 1){
         i = i + 1 # go to next row
         }
       E[i,j] <- ifelse(U[i,j] == -1, U[i,j],-U[i,j]) # case of neg. 1
 
       l[k] <- list(E)  #l is list of elemination matrices
       k = k + 1
       U <- E %*% U
       E <- diag(n)
       i <- i + 1
    }
    j = j + 1
  }
 
  i = length(l)
  L = 0
  X = 0
  while(i >= 1){
    if(i == length(l)) {
      L <- solve(l[[i]])
    } else {
      X <- l[[i]]
      L <- solve(X) %*% L
    }
    i = i -1 
  }
 
  r = list("L" = L, "U" = U, "Z" = L %*% U)
 
  return(r)
}

A <- matrix(c(1,2,3,1,1,3,2,0,1),nrow = 3)

print("test system:")
(f = factormat(A))
f$Z == A




```
### a factorized matrix

$$ 

A = \begin{bmatrix}
1 & 2 & 3 & 4\\
1 & 3 & 2 & 4\\
5 & 4 & 3 & 2\\
4 & 2 & 2& 1 \end{bmatrix}\\
$$

```{r}
A <- matrix(c(1,2,3,4,1,3,2,4,5,4,3,2,4,2,2,1),nrow = 4)

f = (factormat(A))
print("Z = L X U")
f$Z == A

```