---
title: "CBowersoxAssign9"
author: "Cheryl Bowersox"
output: html_document
---

# Week 9 Assignment

Cheryl Bowersox 
IS 605 - 1
3/29/15

## Problem Set 1

### (1) random variable for function:

$$
  f(x) = x , 0 \leq x \leq 1 \\
  f(x) = 2 - x , 1 < x \leq 2 \\
$$  
  CDF:
$$
  F(x) = \frac{x^2}{2}, 0\leq x \leq 1\\
  F(x) = 2x - \frac{x^2}{2} - 1, 1 < x \leq 2\\
$$

$$
  F^{-1}(y) = \sqrt{2y},  0 \leq x \leq .5\\
  F^{-1}(y) = \sqrt{2-2y} + 2,   .5 < x \leq 1\\
$$
```{r}
library(ggplot2)
invf <- function(y) {
  f.y = NULL
  if(y >= 0 && y <=.5){
    f.y <- sqrt(2*y)
  }else{
      if(y <= 1 && y > .5){
        f.y <- 2- sqrt(2-2*y)
      }
    }
  return(f.y)
}

rval <- function(){
  val <- sapply(runif(1),invf)
  return(val)
}

```
Random 5 values following distribution f(x):
`r rval()`
`r rval()`
`r rval()`
`r rval()`
`r rval()`


### (2) random variable for function:
$$
  f(x) = 1-x , 0 \leq x \leq 1\\
  f(x) = x-1 , 1 < x \leq 2
$$

```{r}
px2 <- function(x) {
  #probablity function
  p = NULL
  if(x >= 0 && x <=1){
    p <- 1 -x
  }else{
    if(x <= 2 && x >1){
        p <- x - 1}
  }
  return(p)
}

sampf <- function(n){
  # in is the sample size to return
  vals <- seq(0,2, by = .01)  #values in range 0 to 2
  p <- sapply(vals,px2)  #prob of vals
  ret<- sample(vals, n, replace = TRUE, prob = p)
  return(ret)
}
```


### (3) Plot histogram for both distributions (1) and (2)

Density for 1000 samples of function (1)
```{r}
df <- data.frame(x = replicate(1000,rval()))
ggplot(data = df, aes(x=x)) +
    geom_histogram(aes(y = ..density..), binwidth = .1,
                   fill = "darkcyan", alpha = 0.4)


```

Density for 1000 samples of function (2)
```{r}
l <- sampf(1000)
df <- data.frame(x = sampf(1000))
ggplot(data = df, aes(x=x)) +
    geom_histogram(aes(y = ..density..), binwidth = .1,
                   fill = "darkcyan", alpha = 0.4)

```


### (4) Function to sample and plot means of sample size 'n' and given pdf function
```{r}

plt.mean <- function(n,pdf, lb = 0, ub = 2){
  # n is the sample size
  # pdf is the name of a pdf function that returns the probablity of given x
  # lb is the lower bound of samples to use in this pdf, default 0
  # ub is the upper bound of samples to use in this pdf, default 2
  p <- match.fun(pdf) #p is pdf function
  mean.p <-c()
  i <- 0
  while(i <= 1000){
    vals <- seq(lb,ub, by = .01)  #values in range given by lb, ub
    s <- sample(vals, n, replace = TRUE, prob = sapply(vals,p))
    mean.p <- c(mean.p, mean(s)) # mean of sample
    i <- i + 1
  }
  dfx <- data.frame(mean.p)
  m <- mean(mean.p)
  stv <- sd(mean.p)
  ggplot(data = dfx, aes(x=mean.p)) + 
    geom_histogram(aes(y = ..density..), binwidth = .005, 
                   fill = "darkcyan", alpha = 0.4)+
    stat_function(geom="line", fun=dnorm, arg=list(mean=m, sd=stv),  
                  colour = "#003333")
}

# function for problem 1
px1 <- function(x){
  p = NULL
  if(x >= 0 && x <=1){
    p <- x
  }else{
    if(x <= 2 && x >1){
        p <- 2-x}
  }
  return(p)
}
  
```


### (5) Plot means of (1), (2), compare to normal distribution

Plot means of function (1), with n = 10, 20, 100 with normal curve
```{r}
plt.mean(10,"px1")
plt.mean(20,"px1")
plt.mean(100,"px1")
```

```

Plot means of function (2), with n = 10, 20, 100 with normal curve
```{r}
plt.mean(10,"px2")
plt.mean(20,"px2")
plt.mean(100,"px2")
```

