---
title: "CBowersoxAssign8"
author: "Cheryl Bowersox"
date: "Saturday, March 14, 2015"
output: html_document
---
# Week 8 Assignment

Cheryl Bowersox 
IS 605 - 1
3/22/15

## Problem Set 1 - 

### probablity collegue takes the bus today
Assuming 30 day month:

P(Bus) = $\frac{2}{3}$

P(Train) = $\frac{1}{3}$

P(OnTime|Bus) = $\frac{1}{2}$

P(OnTime|Train) = $\frac{9}{19}$

P(OnTime) = P(OnTime|Bus) + P(OnTime|NotBus)

P(OnTime) = $\frac{2}{3}$ P(OnTime|Bus) + $\frac{1}{3}$ P(OnTime|Train) = $\frac{19}{30}$

P(Bus|OnTime) = $\frac{P(OnTime|Bus) \times P(Bus)}{P(OnTime)}$

P(Bus|OnTime) = $\frac{1}{2} \times \frac{2}{3} \times \frac{30}{19}$

P(Bus|OnTime) = $\frac{10}{19}$ or `r 10/19`


## Problem Set Two

### Build network
```{r}
library(gRain)

lh <- c("low", "high") # simple high/low for most
ny <- c("no", "yes")


d <- cptable(~diff, values = c(.3,.7), levels = ny) # difficult course - Yes/No
i <- cptable(~intel, values = c(.8,.2), levels = lh) #intellegence - high/low
g.i <- cptable(~grade|diff+intel, values = c(.6,.4,.8,.2, .01, .99, .1, .9), levels = lh) 
s.i <- cptable(~sat|intel, values = c(.9, .1, .2, .8), levels=lh)
l.g <- cptable(~letter|grade, values = c(.9, .1, .05, .95), levels = lh)

plist <- compileCPT(list(d, i, g.i, s.i, l.g))
plist$d

net1 <-grain(plist)


```

### Probablity of difficulty when letter was good (high)

```{r}
#Current difficulty
querygrain(net1, nodes=c("diff"), type="marginal")

#New evidence
net12 <- setEvidence(net1, nslist=list(letter = "high"))
#New prob. difficulty 
querygrain(net12, nodes=c("diff"), type="marginal")
````
Probability of difficulty = yes decreases from 70% to about 63% when the additional evidence of a positive letter is introduced.

```{r}
#also Good SAT scores
#New evidence
net13 <- setEvidence(net1, nslist=list(letter = "high", sat = "high"))
#New prob. difficulty 
querygrain(net13, nodes=c("diff"), type="marginal")
```
When additional evidence of a high SAT score, along with positive letter, is introduced, the probability that the difficulty = yes is around 67%. 


