# Week 3 Assignment
Cheryl Bowersox 
IS 605 - 1
2/15/15

========================================================
## Problem Set 1

### 1 Rank of matrix A

$$
A = \begin{bmatrix}
1 & 2 & 3 & 4\\
-1 & 0 & 1 & 3\\
0 & 1 & -2 & 1\\
5 & 4 & -2& 3\end{bmatrix}\\
$$

```{r}
solvmat <- function(A){

  i <- 1
  j <- 1
  n <- nrow(A)
  c <- ncol(A)
  
  while (i <= n ){
  
    # is entire col zero
      if(all(A[,j] == 0)){
        j = j +1
      }
      
      if(A[i,j] == 0){
        #find a non-zero
        w = which(A[,j] == abs(max(A[,j])))
        #swap rows
        s <- A[w,]
        A[w,] <- A[i,]
        A[i,] <- s
      }
      
      # pivot  = 1
      p <- A[i,j] # pivot
      A[i,] <- A[i,]/p
      
      #transform other rows
      k <- 1
      
      while (k <= n){
        if (k != i){
          A[k,] = A[k,] - A[i,] * A[k,j]
        }
        k <- k + 1
      }
    i <- i + 1
    j <- j + 1
    }
  
  
  B <- A
  return(B)
}
```
The rank of matrix A is equal to the number of non-zero rows in reduced echelon form

```{r}
A <- matrix(c(1, -1, 0, 5, 2, 0, 1, 4, 3, 1, -2, -2, 4, 3, 1, -3),nrow = 4)
T <-  solvmat(A)


i<- 1
rank <- 0
while (i <= nrow(A)){
  if(!(all(A[i,] == 0))){ rank = rank + 1}
  i <- i+1
  }
```
In this case the reduced for of A is equal to
```{r}
T
```
which has `r rank` non-zero rows and so has a rank of `r rank` 


### 2 - min and max rank of an nXm matrix where m > n
  (1) matrix A is n x m with m > n
  
  (2) the maximum number of linearly independent rows of A is the minumum of (n, m), in this case n
  
  (3) If  n < m and A is non-zero then A must have at least 1 independent row
  
  (4) the rank of A is between 1 and n

### 3 - the rank of matix B

$$
B = \begin{bmatrix}
1 & 2 & 1\\
3 & 6 & 3\\
2 & 4 & 2\end{bmatrix}\\
$$

(1) The vectors $B{i1}$ and $B{i3}$ are equal and are linearly dependent: $B{i1} = 1 \times B{i3}$

(2) The vectors $B{i1}$ and $B{i2}$ are linearly dependent: $B{i1} = 2 \times B{i3}$

(3) The rank of $B$ is 1

========================================================
## Problem Set 2

### find the eigenvalues of A

$$
A = \begin{bmatrix}
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6 \end{bmatrix}\\
$$

$$
\begin{align}
A\vec{X} = \lambda\vec{X} \\
\\
(\lambda I{3} - A )\vec{X}  = 0 \\
\\
B = \lambda I{3} - A \\
\end{align}
$$

$$
B = \begin{bmatrix}
(\lambda - 1) & 2 & 3\\
0 & (\lambda - 4) & 5\\
0 & 0 & (\lambda - 6) \end{bmatrix}\\
$$

det(B):
$$
\begin{vmatrix}
(\lambda - 1) & 2 & 3\\
0 & (\lambda - 4) & 5\\
0 & 0 & (\lambda -4) \end{vmatrix}  = 0 \\
$$

Characteristic polynomial is:


  $(\lambda - 1) \times [(\lambda - 4)(\lambda - 6) - 0)]  = 0$

  $\lambda^3 - 11\lambda^2 + 34\lambda - 24 = 0$
  $\lambda = {1, 4, 6}$

$$

$\lambda$ = 1

$$
B \vec{X} = 0\\
$$

$$
B = \begin{bmatrix}
(1 - 1) & -2 & -3\\
0 & (1 - 4) & -5\\
0 & 0 & (1 - 6) \end{bmatrix}\\
$$

$$
\begin{bmatrix}
0 & -2 & 3\\
0 & -3 & -5\\
0 & 0 & -5 \end{bmatrix}\\
$$

$$
\begin{align}
-2x_{2} - 3x_{3} = 0 \\
-3x_{2} - 5x_{3} = 0 \\
-5x_{3} = 0 \\

x_{3} = 0 \\
x_{2} = 0 \\
x_{1} = free \\
\end{align}
$$

$$
\vec{X} = \begin{bmatrix}
  x\\
  0 \\
  0 \end{bmatrix} = c
\begin{bmatrix}
  1\\
  0 \\
  0 \end{bmatrix}   for \lambda = 1\\
$$

$\lambda$ = 4

$$
B \vec{X} = 0\\
$$

$$
B = \begin{bmatrix}
(4 - 1) & -2 & -3\\
0 & (4 - 4) & -5\\
0 & 0 & (4 - 6) \end{bmatrix}\\
$$

$$
\begin{bmatrix}
3& -2 & -3\\
0 & 0 & -5\\
0 & 0 & -2 \end{bmatrix}\\
$$

$$
\begin{align}
3x_{1} -2x_{2} -3x_{3} = 0 \\
5-x_{3} = 0 \\
-2x_{3} = 0 \\

x_{3} = 0 \\
x_{2} = 3/2x_{1}\\
x_{1} = free \\
\end{align}
$$

$$
\vec{X} = \begin{bmatrix}
  x\\
  3/2x \\
  0 \end{bmatrix}  = c
\begin{bmatrix}
  1\\
  3/2 \\
  0 \end{bmatrix}   for \lambda = 4\\
$$

$\lambda$ = 6

$$
B \vec{X} = 0\\
$$

$$
B = \begin{bmatrix}
(6 - 1) & -2 & -3\\
0 & (6 - 4) & -5\\
0 & 0 & (6 - 6) \end{bmatrix}\\
$$

$$
\begin{bmatrix}
5 & -2 & -3\\
0 & 2 & -5\\
0 & 0 & 0 \end{bmatrix}\\
$$

$$
\begin{align}
5x_{1} -2x_{2} -3x_{3} = 0 \\
2x_{2} - 5x_{3} = 0 \\

x_{2} = 5/2x_{3} \\
x_{1} = 8/5 x_{3}\\


\end{align}
$$

$$
\vec{X} = \begin{bmatrix}
  8/5x\\
  5/2x \\
  x \end{bmatrix}  = c
\begin{bmatrix}
  8/5\\
  5/2 \\
  1\end{bmatrix}   for \lambda = 6\\
$$
```{r}

```