---
title: "CBowersoxAssign6"
author: "Cheryl Bowersox"
date: "Saturday, March 07, 2015"
output: html_document
---
# Week 5 Assignment

Cheryl Bowersox 
IS 605 - 1
3/8/15


## Problem Set 1

### Number of possible outcomes of roll die 3 times
n = 6
possible outcomes (order matters) = $6^3$ = `r 6^3` 

possible outcomes (order does not matter ) = 

$$
  n = 6\\
  r = 3\\
  t = \frac{(6+3-1)!}{3!(6-1)!}\\
  t = 56
$$


### Probablity of sum = 3 when roll dice twice

$$
  n = 6\\
  p(x) = \frac{1}{6}\\
  p(x,y) = \frac{1}{6^2} = \frac{1}{36}\\
  2*p(x,y) = \frac{1}{18}
$$

### Probablity of same birthday
The probablity that any two people out a of population r share a birthday is

p = 1 - probablity that no two people share a birthday

q = probablity no people share


For 25 people:
$$
  n = 365\\
  k = 25\\
  q = {_{25}P_{365}} \times \frac{1}{365^{25}}
$$

```{r}
cperm <- function(n,k){
  choose(n,k) * factorial(k)
  }

d <- cperm(365,25)/(365^25)
```

$$
  p =1 - `r d` = `r 1 - d`
$$


For 50 people: 
$$
  n = 365\\
  k = 50\\
    q = {_{50}P_{365}} \times \frac{1}{365^{50}}
$$
```{r}
d <- cperm(365,50)/365^50
```

$$
  p =1 - `r d` = `r 1 - d`
$$


## Problem Set 2

### read data and calculate word and word pair probablities 

```{r}
library(stringr)

#get doc
library(RCurl)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
fileurl= "https://raw.githubusercontent.com/cherylb/CompMath/master/assign6.sample.txt"
data <- getURL(fileurl)
doc  <-readChar(fileName, file.info(fileName)$size)



singlewrd <- function(w,doc){
  
  #function takes single word and full document data and finds 
  # number of time a word is contained in the doc
  s <- paste("\\<",w,"\\>", sep ="")
  a <- "\\W+"
  cfull <- sum(str_count(doc, s))  #count number of full words
  cpart <- sum(str_count(doc, w)) # count number of times it appears as part of a word
  ctot <- sum(str_count(doc, a)
  
}

```





```{r}

```

