---
title: "CBowersoxAssign5b"
output: html_document
---

# Week 5 Assignment

Cheryl Bowersox 
IS 605 - 1
3/1/15


## Problem Set 1

```{r}
# function for converting matrix format from R
convmat <- function(X){
  M <- paste("\\begin{bmatrix}","\n")
  
  i = 1
  n = nrow(X)
  if(is.null(n)){
    n <- length(X)
    while(i <= n){
       M <- paste(M, paste(round(X[i],4), collapse = " & "),"\\\\ \n")
       i = i + 1
    }
  }else{
    while(i <= n){
       M <- paste(M, paste(round(X[i,],4), collapse = " & "),"\\\\ \n")
       i = i + 1
    }
  }
  M <- paste(M, "\\end{bmatrix}","\n", collapse = "")
  return (M)
}

A <- matrix(c(1,1,1,1,0,1,3,4), nrow = 4)

b <- matrix(c(0,8,8,19), nrow = 4)


```
$$
  A =
  `r convmat(A)`\\ 
$$

$$
  A^TA = 
  `r convmat(t(A) %*% A)`\\
$$

$$
  A^T\vec{b} = 
  `r convmat(t(A) %*% b)`\\  
$$

### system is consistent and solvable for p
```{r}
# using p to solve system 
p <- matrix(c(1,5,13,17), nrow = 4)
x <- solve(t(A)%*%A) %*% t(A) %*% p

e <- round(p - (A %*% x),6)

E <- round(sum(e^2), 4)
```

The least squares solution  for

$$
 \vec{p} = `r convmat(p)`\\
$$


$$
 \hat{x} = 
  `r convmat(x)`\\
$$
  

With error vector $\vec{e}$ of zero:

$$
  \vec{e} = 
  `r convmat(e)`\\
$$

the zero error indicates that $\hat{x} = \vec{x}$, and $\vec{p}$ is consistent


### e is for orthoganol to p  for C(A)
The error vector $\vec{e}$ is the difference between $\vec{b}$ and $\vec{p}$:

$$
  \vec{e} = `r convmat(b - p)`\\
$$

The dot product of $\vec{e}\$ and $\vec{p}$ is `r round(sum(e*p), 6)` 
e is orthogonal to p

```{r}
i <- 1
B <- rep(NA,ncol(A))
while (i <= ncol(A)){
    B[i] <- (round(sum(e * A[,i])))
    i <- i +1
}

```

The dot products of $\vec{e}$ across the columns of A are

`r B`

$\vec{e}$ is orthogonal to columns of A



## Problem Set 2

### reading data into matrix A and b

```{r}
#import data
# Read matrix data from data on github allows others to access data
library(RCurl)
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl")))
fileurl= "https://raw.githubusercontent.com/cherylb/CompMath/master/auto-mpg.data"
data <- getURL(fileurl)
df <- read.table(text = data)

names(df) <- c("displace", "horsepwr", "weight", "acceler", "mpg")

A <- as.matrix(df[1:4])

b <- as.matrix(df[5])
```

### solving for x using least squares

```{r}

leastsq <- function(A,b){
  # function takes values in matrix A and creates least squares solution for matrix b
  # where A is an n x m matrix and b is a n x 1 matrix
  # returns:
  #   X, the best-fit solution
  #   e, the error vector
  #   E, the squared error
  
  X <- solve(t(A)%*%A) %*% t(A) %*% b
  e <- b - (A %*% X)
  E <- sum(e^2)
  sol <- list("X" = X, "e" = e, "E" = E)
  
  return(sol)
}

#rounded for display in equation
ls <- leastsq(A,b)
X <- round(ls$X, 4)
E <- round(ls$E, 4)

#MSE
MSE <- E/nrow(A)
```

Coefficients: 

$$
  \hat{X} = 
  `r convmat(X)`\\
$$


Equation:

$$
  \hat{mpg} = \\
  `r X[1]` \times displacement + \\
  `r X[2]` \times horsepower + \\
  `r X[3]` \times weight + \\
  `r X[4]` \times acceleration\\
$$

Squared Error: `r E`

Mean Square Error: `r MSE`