# Week 10 Assignment
Cheryl Bowersox 
IS 605 - 1
4/4/15

========================================================
## Problem Set 1

### Create Martrix A for 6 site universe
$A_{ij} = \frac{1}{|P_i|}$ if link from i to j exists, 0 otherwise

```{r}

library(igraph)
# function for converting matrix format from R
convmat <- function(X){
  M <- paste("\\begin{bmatrix}","\n")
  
  i = 1
  n = nrow(X)
  if(is.null(n)){
    n <- length(X)
    while(i <= n){
       M <- paste(M, paste(round(X[i],4), collapse = " & "),"\\\\ \n")
       i = i + 1
    }
  }else{
    while(i <= n){
       M <- paste(M, paste(round(X[i,],4), collapse = " & "),"\\\\ \n")
       i = i + 1
    }
  }
  M <- paste(M, "\\end{bmatrix}","\n", collapse = "")
  return (M)
}
```

```{r}
#define A, 6 X 6 matrix 

A = diag(6)
A[,1] = c(0, .5, .5, 0, 0, 0)
A[,2] = c(0, 0, 1, 0, 0,0)
A[,3] = c(.25, .25, 0, 0, .25, .25)
A[,4] = c(0,0,0,0,.5, .5)
A[,5] = c(0, 0, 0, .5, 0, .5)
A[,6] = c(0,0,.5,.5,0, 0)
```
Matrix for graph is

$$
  A =
  `r convmat(A)`\\ 
$$

### Matrix B with decay

$B = d\times A + \frac{1-d}{n}$
d is the probablity a user will use a link on the current page

1-d being the probablity the user will jump to another page

n is the finite probablity of reaching a web page at random

In this case setting d = .85 and n = 6 gives
```{r}
d <-.85
n <- 6
B <- d * A + (1-d)/n
```

$$
  B =
  `r convmat(B)`\\ 
$$

### converging for large number of iterations
k is the number of iterations

r is uniform rank vector

$r = B^k\times r$ for sufficiently large k

setting k to 10, 50, 100 gives:
```{r}
i <- 1
r = rep(1/n, n)  # intial value for r is uniform distribution
i = 1
while(i <= 10){
  r <- B %*% r
  i <- i +1
  }
k10 <- r
 
while(i <= 50){
  r <- B %*% r
  i <- i +1
  }
k50 <- r

while(i <= 100){
  r <- B %*% r
  i <- i +1
  }
k100 <- r
```
when k is 10 

$$
  r =
  `r convmat(k10)`\\ 
$$


when k is 50 

$$
  r =
  `r convmat(k50)`\\ 
$$

when k is 100 

$$
  r =
  `r convmat(k100)`\\ 
$$

r is converging for suffienctly large k:
  
$$
  r50 - r100  =
  `r convmat(k50-k100)`\\ 
$$

### eigenvalues and eigenvectors of B
```{r}
evalB <- eigen(B)$values
evecB <- eigen(B)$vectors
```
1 is an eigenvalue of B:
$$
  `r convmat(evalB)`\\ 
$$

1 is the maximum eigenvalue of B:
$$
  `r max(Re(evalB))`\\ 
$$

Corresponding eigenvector of B is:
$$
  `r convmat(Re(evecB[,1]))`\\ 
$$

 which is scaler multiple of 
$$
  `r convmat(k100)`\\ 
$$



```{r}
evalB1 <- Re(evalB[1])
dif<- Re((evecB[,1]/k100 )[1])

```
$$
  for \lambda = 1\\
  \vec{X}  =  `r convmat(Re(evecB[,1]))` = `r dif` \times `r convmat(k100)`= 
  `r   convmat(dif * k100)`\\
$$

###using igraph and page.rank package

```{r}
g <- graph.formula(1-+2,1+-+3, 3 +-+ 2, 3+-+6, 3-+5, 6 +-+4, 5-+6, 5 +-+ 4)
rnew  <- page.rank(g, directed = TRUE)$vector
rnew <- matrix(rnew[order(names(rnew))])

plot(g)
```

The r vector  calculated by the page.rank function is the same as that calculated using 100 itterations of matrix B

  New r calculated using page.rank =
$$  
  `r convmat(rnew)`\\
$$
  r calculated using itterations of B =
$$
  `r convmat(k100)`\\
$$