---
title: "Assignment 13"
author: "Cheryl Bowersox"
date: "Saturday, May 02, 2015"
output: html_document
---

## compute the derivitave using limits
$f(x) = x^3 + x^2$

```{r}

fofx <- function(a){
  f <- a^2 +a^3
  return(f)
}
limatx <- function(x){
  lim <- c()
  small.h <- c()
  h <- .1
  # limit of f as h--> 0
  while (h > .000000001){
    lim <- c((fofx(x+h)-fofx(x))/h, lim, (fofx(x-h)-fofx(x))/-h) # 0 from right
    small.h <- c(h, small.h, -h)  #  
    h <- h/5
  }
  r <- data.frame(small.h, lim)
  return(r)
}
```
The result of computing
 $$\lim_{x\to0}\frac{f(x-h) - f(x)}{h}$$
for a given x and very small values of h approaching 0
converges to a single value as h gets closer to 0.
The value of that convergence is the derivative f at x, this can be demonstrated both numerically and graphically


An example where x = 2 gives the following results as h approaches 0
```{r}
x <- 2
vals <- limatx(x)
df <- round(vals[1,2],6)
```
Numerically the mean and median of the resulting values converge to a single value as h approaches 0 from right and left
```{r}
summary(vals)
```
Resulting values for small positive h are:
```{r}
head(vals,5)
```

Resulting values for small negative h are:
```{r}
tail(vals,5)
```

The resulting graph of the values demonstrates the convergence at `r df`

```{r, echo=FALSE}
plot(vals)
```

The derivative f'(x) where x = 2 is `r df`


Example for x = 4.2:

Summary of results
```{r}
summary(limatx(4.2))
```

Plot of results

```{r, echo=FALSE}
plot(limatx(4.2))
```

The derivative f'(x) where x = 4.2 is `r round(limatx(4.2)[1,2],6)`


## compute the area under the curve for x = [1,3]

$f(x) = 3x^2 + 4x$
```{r}

fofx2 <- function(a){
  f <- 3 * a^2 + 4 * a
  return(f)
}
limatx2 <- function(a,b,dx){  #where a is begining of interval, dx is interval
  int <- b - a #size of total interval
  lim <- 0
  x <- a
  if(dx < int){
    while ( x < b){
      lim <- fofx2(x + dx) * dx + lim
      x <- x + dx
    }
  }
 return(lim)
}
v <- c(1:6)
deltax <- 1/10^v

i = 1
result <- c()

while(i < 7){
  result <- c(result,limatx2(1,3,deltax[i]))
  i<- i +1
}
conv <- round(result[6],0)
dfint <- data.frame(deltax,result)
```

For very small intervals the area under f(x) converges to `r conv`

```{r}
dfint
```

Plotting the interval and the resulting value gives:
```{r, echo=FALSE}
plot(dfint)
```

## Solving through analysis

### Integration by parts
$$
  F = \int\sin(x)\cos(x)\,dx\\
  f(x) = \sin(x)\\
  f'(x) = \cos(x)\\
  g'(x) = \cos(x)\\
  g(x) = \sin(x)\\
  F = f(x)g(x) - \int f'(x)g(x)\,dx\\
  F = \sin(x)\sin(x) - \int\cos(x)\sin(x)\,dx\\
  F = \int\sin(x)\cos(x)\,dx = sin(x)^2- \int\cos(x)\sin(x)\,dx\\
  2\int\sin(x)\cos(x)\,dx = sin(x)^2 + C\\
  F = \frac{1}{2}\sin(x)^2 + C
$$

### Integration by parts - 2
$$
  F = \int x^2 e^x\,dx\\
  f(x) = x^2\\
  f'(x) = 2x\\
  g'(x) = e^x\\
  g(x) = e^x\\
  F = x^2e^x - \int 2xe^x\,dx\\
  F = x^2e^x - 2xe^x + 2\int e^x\,dx\\
  F = x^2e^x - 2xe^x +2e^x + C\\
  F = e^x(x^2 - 2x + 2) + C\\
$$

### Product of functions
$$ 
 h'(x) =  \frac{d}{dx}(x cos(x))\\
  f(x) = x\\
  g(x) = cos(x)\\
  f'(x) = 1\\
  g'(x) = -sin(x)\\
  h'(x) = cos(x) - x*sin(x)\\
$$

### Chain rule
$$
  h'(x) =  \frac{d}{dx}(e^{x^4})\\
  f(g) = e^g\\
  g(x) = x^4\\
  f'(g) = e^g\\
  g'(x) = 4x^3\\
  h'(x) = f'(g(x))*g'(x)\\
  h'(x) = e^{x^4}*4x^3 = 4x^3e^{x^4}
$$
  
  h'(x) = 
  


